---
title: "Basic EDA"
author: "John Redden"
date: "2024-04-07"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(corrplot)
```
## Fail/Withdraw Basic EDA
```{r}
studentDf <- read.csv("rawdata/studentInfo.csv")

# New variable 'binary_result'
studentDf$binary_result <- ifelse(studentDf$final_result == "Distinction" | studentDf$final_result == "Pass", "Pass", "Fail")
studentDf$pass <- ifelse(studentDf$final_result == "Distinction" | studentDf$final_result == "Pass", "Pass", "Fail")

studentDf$imd_band[studentDf$imd_band == ""] <- NA
studentDf$imd_band[studentDf$imd_band == "10-20"] <- "10-20%"
studentDf <- na.omit(studentDf)

studentDf$code_module <- factor(studentDf$code_module)
studentDf$code_presentation <- factor(studentDf$code_presentation)
studentDf$gender <- factor(studentDf$gender)
studentDf$region <- factor(studentDf$region)
studentDf$highest_education <- factor(studentDf$highest_education)
studentDf$imd_band <- factor(studentDf$imd_band)
studentDf$age_band <- factor(studentDf$age_band)
studentDf$disability <- factor(studentDf$disability)
studentDf$final_result <- factor(studentDf$final_result)
studentDf$binary_result<- factor(studentDf$binary_result)

# Adjust factor levels
studentDf$binary_result <- factor(studentDf$binary_result, levels = c("Pass", "Fail"))
studentDf$final_result <- factor(studentDf$final_result, levels = c("Distinction", "Pass", "Withdrawn", "Fail"))

str(studentDf)
```
## Counts

```{r}
# Count the number of records for each final_result
result_counts <- studentDf %>%
  group_by(final_result) %>%
  summarise(Count = n())
head(result_counts)

pass_counts <- studentDf %>%
  group_by(binary_result) %>%
  summarise(Count = n())
head(pass_counts)

```

# Visualize the Results

```{r}
ggplot(data = result_counts, aes(x = final_result, y = Count, fill = final_result)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = c("Distinction" = "gray", "Pass" = "gray", "Withdrawn" = "orange", "Fail" = "red")) +
  labs(title = "Count of Student Final Results", x = "Final Result", y = "Count") +
  theme_minimal() +
  theme(legend.title = element_blank())
```

##  Combine Distinction/Pass for Comparisons

```{r}
studentDf <- studentDf %>%
  mutate(plot_result = ifelse(final_result == "Withdrawn", "Withdrawn", as.character(pass)))

# Calculate the counts and percentages for each result category
result_counts <- studentDf %>%
  count(plot_result) %>%
  mutate(Percent = paste0(round((n / sum(n)) * 100, 1), "%"))

# Set the factor levels to ensure the order of the bars in the plot
result_counts$plot_result <- factor(result_counts$plot_result, levels = c("Pass", "Withdrawn", "Fail"))

ggplot(data = result_counts, aes(x = plot_result, y = n, fill = plot_result)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste(n, " (", Percent, ")", sep = "")), 
            vjust = -0.3, size = 3.5, position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Pass" = "gray", "Withdrawn" = "orange", "Fail" = "red")) +
  labs(title = "Aggregate Count of Student Results", x = "Result", y = "Count") +
  theme_minimal() +
  theme(legend.title = element_blank())

```

## Visualize Rates for each Module
```{r}
studentDf$combined_result <- ifelse(studentDf$final_result == "Withdrawn", "Withdrawn", 
                                    ifelse(studentDf$pass == "Pass", "Pass", "Fail"))

# Count the number of each result type within each code_module
result_counts <- studentDf %>%
  group_by(code_module, combined_result) %>%
  summarise(Count = n(), .groups = 'drop')

# Set the factor levels for combined_result to ensure the order of the bars
result_counts$combined_result <- factor(result_counts$combined_result, 
                                       levels = c("Pass", "Withdrawn", "Fail"))

# Plot the graph
ggplot(result_counts, aes(x = code_module, y = Count, fill = combined_result)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("Pass" = "gray", "Withdrawn" = "orange", "Fail" = "red")) +
  labs(title = "Comparison of Result Status by Code Module",
       x = "Code Module",
       y = "Count") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "bottom")

```


## Significant Features for "Withdrawn" aggregate Cohort

```{r}
studentDf$withdrawn <- as.factor(ifelse(studentDf$final_result == "Withdrawn", 1, 0))

modelWithdrawn <- glm(withdrawn ~ code_module + gender + region + highest_education + imd_band + age_band + num_of_prev_attempts + studied_credits + disability, 
             family = binomial(link = "logit"), 
             data = studentDf)

summary(modelWithdrawn)

```

**Intercept**: The log odds of a student withdrawing when all other predictor variables are held at their reference levels.

**Code Module**: The baseline course is "AAA", and the coefficients for other modules indicate their difference in log odds of withdrawing compared to "AAA". For instance, "CCC" (1.4690) has a higher log odds of withdrawal, suggesting students in this module are more likely to withdraw compared to those in "AAA".

**Gender (M)**: The coefficient for males (-0.1150) suggests males are slightly less likely to withdraw compared to females (the reference group), all else being equal.

**Region**: Different regions have different impacts on withdrawal likelihood. For example, students from the "Scotland" region (-0.2401) are less likely to withdraw than those from the reference region.

**Highest Education**: "Lower Than A Level" (0.4018) significantly increases the likelihood of withdrawal, suggesting students with lower initial education levels are more prone to withdrawal.

**IMD Band**: Students from higher IMD bands (indicative of less deprivation) are generally less likely to withdraw, with "90-100%" (-0.4156) showing a substantial decrease in the likelihood of withdrawal.

**Studied Credits (0.0084)**: A positive coefficient indicates that with every additional credit studied, the likelihood of withdrawal slightly increases.

**Disability (Y)**: Having a disability (0.3065) increases the likelihood of withdrawal.

## Significant Features for "Fail"

```{r}
studentDf$fail <- as.factor(ifelse(studentDf$final_result == "Fail", 1, 0))

modelFail <- glm(fail ~ code_module + gender + region + highest_education + imd_band + age_band + num_of_prev_attempts + studied_credits + disability, 
             family = binomial(link = "logit"), 
             data = studentDf)

# Summary of the model to see coefficients and their significance
summary(modelFail)
```

**Code Module**: Different modules have varying risks of failure; for instance, code_moduleGGG significantly increases the likelihood of failing, showing that the module's content or structure may have a higher difficulty or different characteristics affecting student performance.

**Gender**: Males are more likely to fail than females.

**Region**: The effect of the region is mixed; some regions like Scotland significantly increase the risk of failing, while others like the East Midlands Region decrease it. This might reflect socioeconomic or educational disparities.

**Highest Education**: Having a lower than A level education significantly increases the likelihood of failing, indicating prior educational attainment's impact on current studies.

**IMD Band**: Lower IMD bands (indicating higher deprivation) are associated with a higher likelihood of failing, especially noticeable in the most deprived bands (imd_band60-70%, imd_band80-90%, and imd_band90-100%).

**Age Band**: Older age bands (age_band35-55) show a significant negative effect on failing, suggesting that younger students might be at a higher risk of failing.

**Number of Previous Attempts**: An interesting positive coefficient suggests that with each additional attempt, the odds of failing increase, possibly indicating accumulating difficulties or challenges in overcoming academic hurdles.

**studied Credits**: Interestingly, more studied credits are associated with a slightly lower likelihood of failing, contrary to what one might expect.


## Associations

```{r}

# Remove the binary_result and final_result columns 
factor_df <- studentDf %>% select_if(is.factor) %>% 
  select(-binary_result, -final_result)

# Calculate Cramer's V
cramers_v <- function(tab) {
  n <- sum(tab)  # Total observations
  phi_sq <- chisq.test(tab)$statistic
  min_dim <- min(nrow(tab)-1, ncol(tab)-1)
  return(sqrt(phi_sq / (n * min_dim)))
}

results <- matrix(NA, ncol = length(factor_df), nrow = length(factor_df),
                  dimnames = list(names(factor_df), names(factor_df)))

# Loop over all pairs of factor columns and calculate Cramer's V
for (i in 1:(length(factor_df)-1)) {
  for (j in (i+1):length(factor_df)) {
    
    tab <- table(factor_df[, i], factor_df[, j])
    
    results[i, j] <- cramers_v(tab)
    results[j, i] <- results[i, j]
  }
}

results_df <- as.data.frame(as.table(results))

ggplot(results_df, aes(Var1, Var2, fill = Freq)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0.1, limit = c(0,1), space = "Lab", 
                       name="Cramer's V") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(x = '', y = '', title = 'Heatmap of Cramer\'s V (Excluding binary_result and final_result)')


```

---
