---
title: "Demographics+VLE Models"
author: "John Redden"
date: "2024-04-20"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(stats)
library(caret)
library(randomForest)
library(class)  # for KNN
library(car)
```

## Load in the TWO Datasets

Load in the studentInfo.csv demographic data and clean the imd_band.

```{r, message=FALSE}
studentDf <- read.csv("rawdata/studentInfo.csv")

studentDf$pass <- ifelse(studentDf$final_result == "Distinction" | studentDf$final_result == "Pass", "Pass", "Fail")

studentDf$imd_band[studentDf$imd_band == ""] <- NA
studentDf$imd_band[studentDf$imd_band == "10-20"] <- "10-20%"
studentDf <- na.omit(studentDf)

studentDf$code_module <- factor(studentDf$code_module)
studentDf$code_presentation <- factor(studentDf$code_presentation)
studentDf$gender <- factor(studentDf$gender)
studentDf$region <- factor(studentDf$region)
studentDf$highest_education <- factor(studentDf$highest_education)
studentDf$imd_band <- factor(studentDf$imd_band)
studentDf$age_band <- factor(studentDf$age_band)
studentDf$disability <- factor(studentDf$disability)
studentDf$final_result <- factor(studentDf$final_result)
studentDf$pass<- factor(studentDf$pass, levels = c("Pass", "Fail"))

studentDf$studied_credits <- scale(studentDf$studied_credits)
studentDf$num_of_prev_attempts <- scale(studentDf$num_of_prev_attempts)

str(studentDf)
```

Load in new VLE data crated by Richard Han.

````{r, message=FALSE}
interactionDf <- read.csv("rawdata/vle_full_dataset.csv")

interactionDf$code_module <- as.factor(interactionDf$code_module)
interactionDf$code_presentation<- as.factor(interactionDf$code_presentation)

interactionDf$mean_clicks <- scale(interactionDf$mean_clicks)
interactionDf$total_clicks <- scale(interactionDf$total_clicks)


str(interactionDf)
```

## Now Merge the two datasets.

```{r}
merged_df <- merge(studentDf, interactionDf, by = c("id_student","code_module", "code_presentation"), all = TRUE)
merged_df <- na.omit(merged_df)

str(merged_df)
summary(merged_df)

```

Address those outliers!

```{r}
# Calculate IQR
Q1 <- quantile(merged_df$mean_clicks, 0.25)
Q3 <- quantile(merged_df$mean_clicks, 0.75)
IQR <- Q3 - Q1

upper_bound <- Q3 + 1.5 * IQR

# Filter out outliers
filtered_df <- merged_df[merged_df$mean_clicks <= upper_bound, ]

summary(filtered_df)
```

## Identiy Significant features in merged dataset.

```{r}
logistic_model <- glm(pass ~ . -id_student - final_result.x - final_result.y, 
                      data=filtered_df, family=binomial(link='logit'))

summary(logistic_model)
```

## Test for Multicollinearity amoung these predictors!

```{r}
# Calculate VIF
vif_values <- vif(logistic_model)
print(vif_values)

# check for VIF values greater than 10
high_vif <- vif_values[vif_values > 10]
print(high_vif)
```
## Logistic Regression with Significant features only w/cross validation

Identified significant features without muticolinearity: code_module + highest_education + imd_band + 
                  zero_day_ratio + total_clicks

```{r}
set.seed(42)
cv_control <- trainControl(method = "cv", 
                           number = 5,  # number of k-folds for cv
                           summaryFunction = defaultSummary)

cv_model <- train(pass ~ code_module + highest_education + imd_band + 
                  zero_day_ratio + total_clicks,
                  data = filtered_df,
                  method = "glm",
                  family = binomial,
                  trControl = cv_control)

predictions <- predict(cv_model, filtered_df)
conf_matrix <- confusionMatrix(predictions, filtered_df$pass)

print(conf_matrix)
```

## Run CV Random Forest and calculate Accuracy

```{r}
set.seed(42)
cv_control <- trainControl(method = "cv", 
                           number = 5,  # number of k-folds for cv
                           summaryFunction = defaultSummary)

cv_rf_model <- train(pass ~ code_module + highest_education + imd_band + 
                    zero_day_ratio + total_clicks, 
                     data = filtered_df,
                     method = "rf",
                     trControl = cv_control,
                     tuneLength = 1) # number of mtry values

predictions <- predict(cv_rf_model, filtered_df)

conf_matrix <- confusionMatrix(predictions, filtered_df$pass)

print(conf_matrix)
```


## Run KNN with CV

```{r}
set.seed(42)
cv_control <- trainControl(method = "cv",
                           number = 5,
                           savePredictions = "final",
                           summaryFunction = defaultSummary,
                           allowParallel = TRUE)

knn_model <- train(pass ~ code_module + highest_education + imd_band + 
                    zero_day_ratio + total_clicks,
                   data = filtered_df,
                   method = "knn",
                   trControl = cv_control,
                   tuneLength = 1)  # Try different values for 'k'

predictions <- predict(knn_model, filtered_df)

conf_matrix <- confusionMatrix(predictions, filtered_df$pass)
print(conf_matrix)

print(knn_model$results)

```















